Multi-modality Imagery Database for Plant Phenotyping

### 2015-05-02 datapaper_baseline.tex
figure update

\subsection{Performance and analysis}
We apply our algorithm on all $144$ frames of each video and evaluate the performance on labeled $36$ frames.
Leaf alignment is applied to the last frame of each video. 
Figure~\ref{fig:alignResult} shows some examples of leaf alignment results. 
Our framework works very well on segmenting large leaves with no overlap to neighbor leaves. 
For overlapping leaves, it becomes more challenging as the edges between the overlapping area are more difficult to be detected. 
However, when the overlapping leaves are further away from the center, they will have a higher chance to be detected as shown in $(3)$ of Figure~\ref{fig:alignResult}. 
When the overlapping leaves are close to the center, smaller leaves will be covered by larger leaves as shown in $(1), (4), (5)$ of Figure~\ref{fig:alignResult}.

@@

\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/9_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/10_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/11_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/12_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/14_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/15_2}\\
(1) & (2) & (3) & (4) & (5) & (6) \\
\end{tabular}
\caption{Leaf alignment results on the last frame of $6$ plant videos. First row shows the original image. Second row shows the alignment result. }
\label{fig:alignResult}
\end{centering}
\end{figure*}

\begin{figure*}
\begin{centering}
\begin{tabular}{c c@{} c@{} c@{} c@{} c@{} c@{} c@{} c@{} c@{}}
%\begin{tabular}{lllllllll}
(1) &
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_1}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_2}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_3}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_4}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_5}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_6}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_7}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_8}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/1_9}\\
(2)&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_1}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_2}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_3}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_4}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_5}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_6}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_7}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_8}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/2_9}\\
(3) &
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.11\textwidth]{Figures/trackExample/3_9}\\

@@

\begin{figure*}
\centering
\includegraphics[width=.98\textwidth]{Figures/performance}\\
\caption{Performance of the baseline method.}
\label{fig:performance}
\end{figure*}

For quantitative evaluation, we vary $\tau$ from $0$ to $1$ and generate the first three evaluation metrics, as shown in Figure~\ref{fig:performance}.
{\it{ULR}} decreases as $\tau$ increases as more leaves are being considered as matched leaves.  
As $\tau$ keep increasing, {\it{ULR}} approaches a constant value, which is the the diferent number in leaf counting that results from both miss detection and false alarms. 
{\it{LE}} increases as $\tau$ increases as it includes leaves with larger tip-based errors for averaging. 
{\it{TC}} increases as $\tau$ increases as more leaves are being considered as correctly tracked leaves. 
Overall, our method can detect $80\%$ and track $60\%$ of all labeled leaves with $XX$ average tip-based errors.

We generate a label image for each frame based on the leaf segmentation results anc compute the {\it{SBD}} score for each labeled image. 
The average {\it{SBD}} over all images is $0.59$.

>>>>

\subsection{Performance and analysis}
We apply our algorithm on all $144$ frames of each video and evaluate the performance on labeled $36$ frames.
Leaf alignment is applied to the last frame of each video.
Figure~\ref{fig:alignResult} shows some examples of leaf alignment results.
Our framework works very well on segmenting large leaves with no overlap to neighbor leaves.
For overlapping leaves, it becomes more challenging as the edges between the overlapping area are more difficult to be detected.
However, when the overlapping leaves are further away from the center, they will have a higher chance to be detected as shown in $(c)$ of Figure~\ref{fig:alignResult}.
When the overlapping leaves are close to the center, smaller leaves will be covered by larger leaves as shown in $(a), (d), (e)$ of Figure~\ref{fig:alignResult}.

@@

\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/9_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/10_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/11_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/12_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/14_2}&
\includegraphics[width=.14\textwidth]{Figures/AlignPerformance/15_2}\\
(a) & (b) & (c) & (d) & (e) & (f) \\
\end{tabular}
\caption{Leaf alignment results on the last frame of $6$ plant videos. First row shows the original image. Second row shows the alignment result. }
\label{fig:alignResult}
\end{centering}
\end{figure*}

\begin{figure*}
\begin{centering}
\begin{tabular}{c c@{} c@{} c@{} c@{} c@{} c@{} c@{} c@{} c@{}}
%\begin{tabular}{lllllllll}
(a) &
\includegraphics[trim= 10 100 20 100, clip, width=.1\textwidth]{Figures/trackExample/1_1}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_2}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_3}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_4}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_5}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_6}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_7}&
\includegraphics[trim= 50 150 100 200, clip, width=.1\textwidth]{Figures/trackExample/1_8}&
\includegraphics[trim= 0 100 20 100, clip, width=.1\textwidth]{Figures/trackExample/1_9}\\
(b)&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_1}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_2}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_3}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_4}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_5}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_6}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_7}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_8}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/2_9}\\
(c) &
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}&
\includegraphics[width=.1\textwidth]{Figures/trackExample/3_9}\\

@@

\begin{figure*}
\centering
\includegraphics[width=.98\textwidth]{Figures/performance.pdf}\\
\caption{Performance of the baseline method.}
\label{fig:performance}
\end{figure*}

For quantitative evaluation, we vary $\tau$ from $0$ to $1$ and generate the first three evaluation metrics, as shown in Figure~\ref{fig:performance}.
{\it{ULR}} decreases as $\tau$ increases as more leaves are being considered as matched leaves.
As $\tau$ keep increasing, {\it{ULR}} approaches a constant value, which is the the diferent number in leaf counting that results from both miss detection and false alarms.
{\it{LE}} increases as $\tau$ increases as it includes leaves with larger tip-based errors for averaging.
{\it{TC}} increases as $\tau$ increases as more leaves are being considered as correctly tracked leaves.
Overall, our method can detect $87\%$ and track $50\%$ of all labeled leaves with less than $20\%$ average tip-based errors.

We generate a label image for each frame based on the leaf segmentation results anc compute the {\it{SBD}} score for each labeled image.
The average {\it{SBD}} over all images is $0.61$.

### 2015-05-02 datapapr_labeling.tex
baseline
xm's first edits
figure update

We use fluorescence images for labeling because of their simple background.
For Arabidopsis images, we label $4$ frames each day.
While for bean images, we label $7$ frames each day because of their spontaneous and faster leaf movement.
A Matlab-based GUI interface is developed for leaf labeling, as shown in Figure~\ref{fig:label}.
A user can open a plant image to label the two tips and annotate each leaf segment.
The results will be automatically saved once a user moves onto the next image for labeling.
For consistent annotation of the same leaf over time, we show a number on the center of each leaf indicating the order of labeling from the previous frame.
%This GUI is used to annotate leaves in each frame.
%For consistent annotation of the same leaf over time, we use the labeled results of each frame to find the correspondence according to leaf centers.
%Incorrect correspondence will be manually corrected.

@@

The labeling of the leaf segment is implemented by clicking the boundary of one leaf at each time.
The labeled leaf boundary is overlaid on the image for better visualization to guide the next action.
Incorrect label can be deleted right after the labeling.
This process continues until all leaf segments have been annotated.
After the labeling of one plant, we visually go through the results and correct inaccurate labels.
Note that one alternative approach for labeling leaf segments is to directly label the membership of superpixels instead of drawing a polygon along the boundary.
Our experience is that since a noticeable percentage of extracted superpixels cover pixels of two neighboring leaves, the extra effort of breaking a superpixel into two makes it a less efficient alternative.
One example of the labeling results for one plant is shown in Figure~\ref{fig:LabelExample}, where one color is used to represent each specific leaf.
As we can see during the transition between day $4$ and day $5$, there is one leaf showing up and covering up the leaf underneath, which disappears and will not be annotated later.

\subsection{Name conventions and file types}
We release training and testing sets in two separate folders.
In each folder, there are two subfolders named Arabidopsis and Bean.
The files in each subfolder have the following form:
\begin{itemize}
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$rgb.png: the original RGB color images;
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$fmp.png: the original fluorescence images;
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$ir.png: the original IR images;
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$depth.png: the original depth images;
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$label.png: the labeled images of fluorescence modality;
  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$tips.txt: the labeled tip locations;
\end{itemize}
where XX indicates the plant type ("AR" or "BE"), X is an integer indicating the date (e.g., $1-9$ for Arabidopsis), YY represents the image index within a day (e.g., $1-16$ for Arabidopsis), and ZZ is the subject (or plant) ID (e.g., $1-5$ for bean).
For each combination of day and hour, we provide four modalities in PNG files ($\_$rgb, $\_$fmp, $\_$ir, $\_$depth).
For annotated images, we have two additional files ($\_$label, $\_$tips) saving the annotation results.
Leaf segmentation results are encoded as indexed PNG files, where each leaf is assigned a unique and consistent leaf ID over time.
Leaf ID starts from $1$ and continuously increase till the totoal number of leaves.
And the background is encoded as $0$.
Tips locations are saved in TXT files where each line has the following format:
\begin{itemize}
  \item leaf ID \quad tip1$\_$x \quad tip1$\_$y \quad tip2$\_$x \quad  tip2$\_$y
\end{itemize}
where leaf ID is an integer number that is consistent with the segmentation label in the PNG file.
tip1$\_$x and tip1$\_$y represent the coordinates of the outer tip point.
tip2$\_$x and tip2$\_$y represent the coordinates of the inner tip point.

@@

\subsection{Experimental Protocols}

As shown in Table~\ref{tab:database}, MSU-PID can be used for applications such as leaf segmentation, leaf alignment, leaf tracking, and leaf counting.
To facilitate future research, we separate the database into the training set and the testing set.
$40\%$ of the data is used for training and $60\%$ for testing.
Specifically, $6$ plants of Arabidopisis and $2$ plants of bean are selected for training.
We will provide training and testing data in different folders.
For fair comparison, both supervised learning and unsupervised learning methods should evaluate their performances on the training and testing sets separately.
The user may decide to utilize one or multiple modalities of the plant imagery for training and testing respectively.
The availability of multiple modalities allows user to design novel experimental setup.
For example, using RGB and depth modalities for training and RGB for testing can evaluate an algorithm that can handle a missing modality during the testing.
%The evaluation metrics will be discussed in Sec.~\ref{sec:baseline}.

>>>>

We use the fluorescence images as the input for labeling because of their simple and uniform background.
For Arabidopsis images, we label $4$ frames each day.
While for bean images, we label $7$ frames each day because of their spontaneous and faster leaf movement.
A Matlab-based GUI interface is developed for leaf labeling, as shown in Figure~\ref{fig:label}.
A user can open a plant image to label the two tips and annotate each leaf segment.
The results will be automatically saved once a user moves onto the next image for labeling.
For consistent annotation of the same leaf over time, we show a number on the center of each leaf indicating the order of labeling from the previous frame.
%This GUI is used to annotate leaves in each frame.
%For consistent annotation of the same leaf over time, we use the labeled results of each frame to find the correspondence according to leaf centers.
%Incorrect correspondence will be manually corrected.

@@

The labeling of the leaf segment is implemented by clicking the boundary of one leaf at each time.
In order to provide more accuracy labeling, we click very dense points ($\sim20$ points on average) on the boundary. 
The labeled leaf boundary is overlaid on the image for better visualization to guide the next action.
Incorrect label can be deleted right after the labeling.
This process continues until all leaf segments have been annotated.
After the labeling of one plant, we visually go through the results and correct inaccurate labels.
One example of the labeling results for one plant is shown in Figure~\ref{fig:LabelExample}, where one color is used to represent each specific leaf.
As we can see during the transition between day $4$ and day $5$, there is one leaf showing up and covering up the leaf underneath, which disappears and will not be annotated later.
In total, we labeled $5142$ leaves.

Note that one alternative approach for labeling leaf segments is to directly label the membership of superpixels instead of drawing a polygon along the boundary.
Our experience is that since a noticeable percentage of extracted superpixels cover pixels of two neighboring leaves, the extra effort of breaking a superpixel into two makes it a less efficient alternative.

\subsection{Name conventions and file types}
We release training and testing sets in two separate folders.
In each folder, there are two subfolders named Arabidopsis and Bean.
The files in each subfolder have the following form:

\begin{itemize}
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$rgb.png: the original RGB color images;
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$fmp.png: the original fluorescence images;
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$ir.png: the original IR images;
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$depth.png: the original depth images;
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$label.png: the labeled images of fluorescence modality;
  \item plant$\_$ID$\_$day$\_$X$\_$hour$\_$YY$\_$tips.txt: the labeled tip locations;
\end{itemize}
where ID indicates the plant ID number ($1$ to $16$ for Arabidopsis, $1$ to $5$ for bean), X is an integer indicating the date ($1-9$ for Arabidopsis, $1$ to $5$ for bean), and YY represents the image index within a day ($1-16$ for Arabidopsis, $1$ to $14$ for bean).
For each combination of day and hour, we provide four modalities in PNG files ($\_$rgb, $\_$fmp, $\_$ir, $\_$depth).
%\begin{itemize}
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$rgb.png: the original RGB color images;
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$fmp.png: the original fluorescence images;
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$ir.png: the original IR images;
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$depth.png: the original depth images;
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$label.png: the labeled images of fluorescence modality;
%  \item plant$\_$XX$\_$day$\_$X$\_$hour$\_$YY$\_$ID$\_$ZZ$\_$tips.txt: the labeled tip locations;
%\end{itemize}
%where XX indicates the plant type ("AR" or "BE"), X is an integer indicating the date (e.g., $1-9$ for Arabidopsis), YY represents the image index within a day (e.g., $1-16$ for Arabidopsis), and ZZ is the subject (or plant) ID (e.g., $1-5$ for bean).
%For each combination of day and hour, we provide four modalities in PNG files ($\_$rgb, $\_$fmp, $\_$ir, $\_$depth).
For annotated images, we have two additional files ($\_$label, $\_$tips) saving the annotation results.
Leaf segmentation results are encoded as indexed PNG files, where each leaf is assigned a unique and consistent leaf ID over time.
Leaf ID starts from $1$ and continuously increase till the total number of leaves.
And the background is encoded as $0$.
Tips locations are saved in TXT files where each line has the following format:
\begin{itemize}
  \item leaf ID \quad tip1$\_$x \quad tip1$\_$y \quad tip2$\_$x \quad  tip2$\_$y
\end{itemize}
where leaf ID is an integer number that is consistent with the segmentation label in the PNG file.
tip1$\_$x and tip1$\_$y represent the coordinates of the outer tip point.
tip2$\_$x and tip2$\_$y represent the coordinates of the inner tip point.

@@

\subsection{Experimental Protocols}
As shown in Table~\ref{tab:database}, MSU-PID can be used for applications such as leaf segmentation, leaf alignment, leaf tracking, and leaf counting.
To facilitate future research, we separate the database into the training set and the testing set.
$40\%$ of the data is used for training and $60\%$ for testing.
Specifically, $6$ plants of Arabidopisis and $2$ plants of bean are selected for training.
%We will provide training and testing data in different folders.
For fair comparison, both supervised learning and unsupervised learning methods should evaluate their performances on the training and testing sets separately.
The user may decide to utilize one or multiple modalities of the plant imagery for training and testing respectively.
The availability of multiple modalities allows user to design novel experimental setup.
For example, using RGB and depth modalities for training and RGB for testing can evaluate an algorithm that can handle a missing modality during the testing.
%The evaluation metrics will be discussed in Sec.~\ref{sec:baseline}.
